---
title: "Customer Behavior"
author: "Tom_Kinyanjui_Njoroge"
date: "July 18, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Defining the Question
1. To Perform clustering from our analysis and visualizations.

2. Apply K-Means clustering vs Hierarchical clustering 


###The metric of Success
Understand the characteristics of Kira Plastinina Customers and apply unsupervised machine learning algorithms 


## Reading the data 
```{r}
shoppers <- read.csv("file:///C:/Users/HP/Downloads/online_shoppers_intention.csv")
head(shoppers)
```
## Check the Data
```{r}
# check the structure of Data
str(shoppers)
```

```{r}
dim(shoppers)
```

```{r}
summary(shoppers)
```

```{r}
 # convert into a factor
shoppers$VisitorType <- factor(shoppers$VisitorType)
head(shoppers$VisitorType)
shoppers$Weekend <- factor(shoppers$Weekend)
head(shoppers$Weekend)
shoppers$Revenue <- factor(shoppers$Revenue)
head(shoppers$Revenue)
shoppers$Month <- factor(shoppers$Month)
head(shoppers$Month)
```

## Data Cleaning
```{r}
#check for missing values
colSums(is.na(shoppers))
anyNA(shoppers)
mice_mod <- mice(shoppers[, c("Administrative","Administrative_Duration","Informational","Informational_Duration","ProductRelated","ProductRelated","ProductRelated_Duration","BounceRates","ExitRates")], method='rf')
mice_complete <- complete(mice_mod)
shoppers$Administrative <- mice_complete$Administrative
shoppers$Administrative_Duration  <- mice_complete$Administrative_Duration
shoppers$Informational <- mice_complete$Informational
shoppers$Informational_Duration <- mice_complete$Informational_Duration
shoppers$ProductRelated <- mice_complete$ProductRelated
shoppers$ProductRelated_Duration <- mice_complete$ProductRelated_Duration
shoppers$BounceRates <- mice_complete$BounceRates
shoppers$ExitRates <- mice_complete$ExitRates
```


```{r}
#checking for duplicates 
shoppers <- unique(shoppers)
dim(shoppers)
```
###checking for outliers 
```{r}
num_col <- shoppers[ ,c(1,2,3,4,5,6,7,8,9,10,12,13,14,15)]
outlier_detection = function(x){
  for(i in colnames(x)){
    boxplot(shoppers[[i]], xlab=i, main=paste0("Boxplot for ",i))
  }
}
outlier_detection(num_col)
```
###dealing with outliers 
```{r}
# replace outliers with the 5th and 95th percentile
outlier_replace <- function(x){
   qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
   caps <- quantile(x, probs=c(.05, .95), na.rm = T)
   H <- 1.5 * IQR(x, na.rm = T)
   x[x < (qnt[1] - H)] <- caps[1]
   x[x > (qnt[2] + H)] <- caps[2]
   return(x)
}
shoppers$Administrative <- outlier_replace(shoppers$Administrative)
shoppers$Administrative_Duration <-outlier_replace(shoppers$Administrative_Duration)
shoppers$Informational <- outlier_replace(shoppers$Informational)
shoppers$Informational_Duration <- outlier_replace(shoppers$Informational_Duration )
shoppers$ProductRelated <- outlier_replace(shoppers$ProductRelated)
shoppers$ProductRelated_Duration <- outlier_replace(shoppers$ProductRelated_Duration)
shoppers$BounceRates <- outlier_replace(shoppers$BounceRates)
shoppers$ExitRates <- outlier_replace(shoppers$ExitRates)
shoppers$PageValues <- outlier_replace(shoppers$PageValues)
shoppers$SpecialDay <- outlier_replace(shoppers$SpecialDay)
shoppers$OperatingSystems <- outlier_replace(shoppers$OperatingSystems)
shoppers$Browser <- outlier_replace(shoppers$Browser)
shoppers$Region <- outlier_replace(shoppers$Region)
shoppers$TrafficType <- outlier_replace(shoppers$TrafficType)
# check to see if there are more outliers
outlier_detection(num_col)
```

```{r eval=FALSE, include=FALSE}
# replace the remaining outliers with NA then predict the values
outlier_na = function(x){
  col <- c(2,3,4,5,6,7,8,9,10,13,15)
  for (i in col){
     outlier <- boxplot.stats(x)$out
     shoppers[x %in% outlier, names(shoppers)[i]] <- NA 
  }
}
shoppers$Administrative_Duration <-outlier_na(shoppers$Administrative_Duration)
shoppers$Informational <- outlier_na(shoppers$Informational)
shoppers$Informational_Duration <- outlier_na(shoppers$Informational_Duration )
shoppers$ProductRelated <- outlier_na(shoppers$ProductRelated)
shoppers$ProductRelated_Duration <- outlier_na(shoppers$ProductRelated_Duration)
shoppers$BounceRates <- outlier_na(shoppers$BounceRates)
shoppers$ExitRates <- outlier_na(shoppers$ExitRates)
shoppers$PageValues <- outlier_na(shoppers$PageValues)
shoppers$SpecialDay <- outlier_na(shoppers$SpecialDay)
shoppers$Browser <- outlier_na(shoppers$Browser)
shoppers$Region <- outlier_na(shoppers$Region)
shoppers$TrafficType <- outlier_na(shoppers$TrafficType)
colSums(is.na())
head(shoppers)

##EDA 

### Univariate
#### Categorical Variables

library("ggplot2" )
# Group Special Days
specialday <- ggplot(shoppers, aes(x=factor(SpecialDay), fill = factor(SpecialDay))) + geom_bar()
specialday +  scale_fill_discrete(name = "Special Day", labels = c("Not Special","Special"))+ labs(title="Number of Special Days", x="Special Days")
# Count the Months
months <- ggplot(shoppers ,aes(x=Month , fill=factor(Month))) + geom_bar() + labs(title = "Distribution of the Months", x="Months")
months +scale_fill_discrete(name = "Month") 
# Count on Visitor Type
visitor <- ggplot(shoppers, aes(VisitorType, fill=factor(VisitorType)))+ geom_bar() + labs(title = "Distribution of Visitor Types", x="Special Days")
visitor + scale_fill_discrete(name = "Visitor Type") 
# Count on Revenue
revenue <- ggplot(shoppers, aes(Revenue, fill=factor(Revenue))) +geom_bar() + labs(title = "Distribution of Revenue")
revenue +scale_fill_discrete(name = "Revenue") 
# Group Revenue by Months
revenue1 <- ggplot(shoppers, aes(x=Month, fill= factor(Revenue)))+ geom_bar(position=position_dodge2(width = 1.2, preserve = "single"))
revenue1 + labs(title = "Distribution of Revenue in a Month") +scale_fill_discrete(name = "Revenue") 
# Group Special Days by Month
specialday1 <- ggplot(shoppers, aes(x=Month, fill= factor(SpecialDay)))+ geom_bar(position = position_dodge2(width = 1.2, preserve = "single"))
specialday1 + scale_fill_discrete(name = "Special Day", labels = c("Not Special","Special")) + labs(title = "Distribution of Special Days in a Month")
# Group Visitor Type by Month
visitor1 <- ggplot(shoppers, aes(x=Month, fill=factor(VisitorType)))+geom_bar(position=position_dodge2(width = 1.2, preserve = "single")) + labs(title = "Distribution of Visitor Type in a Month")
visitor1 + scale_fill_discrete(name = "Visitor Type") 
# Group Weekend by Month
weekend <- ggplot(shoppers, aes(x=Month, fill=factor(Weekend)))+geom_bar(position=position_dodge2(width = 1.2, preserve = "single"))+labs(title="Distribution of Revenue during weekends over the Months",subtitle = "(FALSE -No Revenue vs TRUE - Revenue)", x ="")+facet_wrap(~ factor(Revenue) )
weekend + scale_fill_discrete(name = "Weekend", labels = c("Not Weekend","Weekend")) 
# Distribution of Traffic Type
traffictype <- ggplot(shoppers, aes(x=factor(TrafficType), fill=factor(TrafficType)))+ geom_bar()+labs(title="Distribution of Traffic Type", x= "Traffic Type")
traffictype +scale_fill_discrete(name = "Traffic Type")
# Distribution of Browser
browser <-ggplot(shoppers, aes(x=factor(Browser), fill=factor(Browser)))+ geom_bar()+labs(title="Distribution of Browser", x = "Browser")
browser +scale_fill_discrete(name = "Browser")
# Distribution of Operating System
os <- ggplot(shoppers, aes(x=factor(OperatingSystems), fill= factor(OperatingSystems)))+ geom_bar()+labs(title="Distribution of  Operating Systems" , x="Operating Systems")
os +scale_fill_discrete(name = "Operating Systems") 
# Distribution of TRaffic Type in a Month
traffictype1 <- ggplot(shoppers, aes(x=Month, fill=factor(TrafficType))) +geom_bar(width = 0.8,position=position_dodge(width = 0.8))+labs(title="Distribution of Traffic Type in a Month")
traffictype1 + scale_fill_brewer(name ="Traffic Type",palette="Set1")
```

```{r Histograms}
histogram = function(x){
  for(i in colnames(x)){
    hist(shoppers[[i]], breaks = 10,main =i,xlab = i,col = "dodgerblue")
  }
}
histogram(num_col)
```
### Bivariate
#### Correlation
```{r}
# Convert the Revenue Col to Numeric
shoppers$Revenue <- as.numeric(shoppers$Revenue)
nums <- shoppers[ ,c(1,2,3,4,5,6,7,8,9,10,12,13,14,15,18)]
# Correlation matrix
corr_matrix = cor(nums)
corr <- as.data.frame(round(corr_matrix,2))
corr
```

```{r Scatter Plot}
admin <- shoppers$Administrative
admin_d <- shoppers$Administrative_Duration
info <- shoppers$Informational
info_d <- shoppers$Informational_Duration
prod <- shoppers$ProductRelated
prod_d <- shoppers$ProductRelated_Duration
exit <- shoppers$ExitRates
bounce <- shoppers$BounceRates
page <- shoppers$PageValues
# Administrative
plot(admin,exit,main = "Relationship between the Administrative and Exit Rate", xlab="Administrative", ylab="Exit Rate")
plot(admin,bounce,main = "Relationship between the Admintrative and Bounce Rate",xlab="Administrative",ylab = "Bounce Rate")
plot(admin,page, main = "Relationship between the Admintrative and Page Values",xlab="Administrative",ylab = "Page Value")
plot(admin, admin_d = "Relationship between the Administrative and Administrative Duration",xlab="Administrative",ylab = "Administrative Duration")
# Informational
plot(info,exit,main = "Relationship between the Informational and Exit Rate",xlab = "Informational",ylab = "Exit Rates")
plot(info,bounce,main = "Relationship between the Informational and Bounce Rate",xlab = "Informational", ylab = "Bounce Rates")
plot(info,page, main = "Relationship between the Informational and Page Values",xlab = "Informational",ylab = "Page Values")
plot(info, info_d, main = "Relationship between the Informational and Informational Duration", xlab = "Informational",ylab = "Informational Duration")
# Product Related
plot(prod,exit,main = "Relationship between the Product Related and Exit Rate",xlab="Product Related",ylab = "Exit Rates")
plot(prod,bounce,main = "Relationship between the Product Related and Bounce Rate",xlab="Product Related",ylab = "Bounce Rates")
plot(prod,page, main = "Relationship between the Product Related and Page Values",xlab="Product Related",ylab = "Page Values")
plot(prod, prod_d, main = "Relationship between the Product Related and ProductRelated Duration",xlab="Product Related",ylab ="Product Related Duration")
#Duration
plot(admin_d,info_d, main = "Relationship between Admin Duration and Informational Duration", xlab = "Administrative Duration", ylab = "Informational Duration")
plot(prod_d, info_d,main = "Relationship between ProductRelated Duration and Informational Duration", xlab = "Product Related Duration",ylab = "Informational Duration")
plot(admin_d,prod_d,main = "Relationship between Admin Duration and ProductRelated Duration", xlab = "Administrative Duration", ylab = "Product Related Duration")
```


##Solution

### Hierarchial Clustering
```{r Hclust}
# Scale our data first
data <- shoppers[,-18]
data <- scale(data)
head(data)
#calculate the Euclidean distance
d <- dist(data, method = "euclidean")
# apply the h-clustering and use different methods
model_comp <- hclust(d, method = "complete")
model_ward <- hclust(d, method = "ward.D")
model_ward2 <- hclust(d, method = "ward.D2")
model_sing <- hclust(d, method = "single")
model_avg <- hclust(d, method = "average")
model_mc <- hclust(d, method = "mcquitty")
model_med <- hclust(d, method = "median")
model_cent <- hclust(d , method = "centroid")
# Plot Dendograms
plot(model_comp, cex = 0.6, hang = -1)
plot(model_ward, cex = 0.6, hang = -1)
plot(model_ward2, cex = 0.6, hang = -1)
plot(model_sing, cex = 0.6, hang = -1)
plot(model_avg, cex = 0.6, hang = -1)
plot(model_mc, cex = 0.6, hang = -1)
plot(model_med, cex = 0.6, hang = -1)
plot(model_cent, cex = 0.6, hang =-1)
```
### K-Means Clustering
```{r}
# Transform Factors to Numeric
shoppers$Month <- as.numeric(shoppers$Month)
shoppers$VisitorType <- as.numeric(shoppers$VisitorType)
shoppers$Weekend <- as.numeric(shoppers$Weekend)
str(shoppers)
```

```{r}
# K-Means is affected by Imbalanced data
table(shoppers$Revenue)
prop.table(table(shoppers$Revenue))
# our class label is imbalanced thus we shall balance it
# Split to train and test
library(ROSE)
library(caret)
# Make the Revenue Class a factor again
shoppers$Revenue <- as.factor(shoppers$Revenue)
# Use ROSE, it helps us to generate data synthetically as well
train.rose <- ROSE(Revenue ~ ., data = shoppers, seed = 1)$data
table(train.rose$Revenue)
feature <- train.rose[, -18]
label <- train.rose[,"Revenue"]
```

```{r}
# Normalize the data
normalize <- function(x){
  return ((x-min(x)) / (max(x)-min(x)))
}
# normalize the train
feat_norm <- as.data.frame(lapply(feature, normalize))
summary(feat_norm)
```

```{r Kmeans}
#Apply the model
model <- kmeans(feat_norm,2, nstart = 30)
# check for no. of records in each cluster
model$size
# a set of initial (distinct) cluster centers.
model$centers
# shows the cluster where each record falls
#model$cluster
#Verfy the results
plot(feat_norm[c(1:2)], col = model$cluster)
plot(feat_norm[c(1:2)], col = label)
plot(feat_norm[c(3:4)], col = model$cluster)
plot(feat_norm[c(3:4)], col = label)
plot(feat_norm[c(5:6)], col = model$cluster)
plot(feat_norm[c(5:6)], col = label)
```

```{r Accuracy }
tb <- table(model$cluster, label)
tb
accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}
paste0("Accuracy: ",accuracy(tb))
```

Both models performed well with accuray above 75% which is optimal 